
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project1

Package used:
```{r dplyr, echo=TRUE, results='hide'}
library(dplyr)
```

## Loading and preprocessing the data

```{r load}
data <- read.csv("./activity.csv", header = TRUE)
head(data)
```

```{r}
str(data)
```

```{r}
summary(data)
```

-------------------
## What is mean total number of steps taken per day?

  Calculating and reporting the total number of steps taken per day. The red line on the plot is the mean of steps taken per day.

```{r}
perday <- group_by(data, date)
stepsperday <- summarise(perday, steps = sum(steps, na.rm =TRUE))
head(stepsperday)
barplot(stepsperday$steps, names = stepsperday$date, xlab = "Date", ylab ="Number of steps", main ="Total number of steps taken per day")
abline(h=mean(stepsperday$steps), col="red", lwd=2)
```

  Calculating and reporting the mean of the total number of steps taken per day
```{r}
meanstepsperday <- mean(stepsperday$steps)
meanstepsperday
```

 Calculating and reporting the median of the total number of steps taken per day
```{r}
medianstepsperday <- median(stepsperday$steps)
medianstepsperday
```

## What is the average daily activity pattern?

  Making a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
perinterval <- group_by(data, interval)
stepsperinterval <- summarise(perinterval, steps = mean(steps, na.rm =TRUE))
head(stepsperinterval,3)
tail(stepsperinterval,3)

plot(stepsperinterval$interval, stepsperinterval$steps, type = "l", xlab = "5-minute interval", ylab = "Average number of steps taken, averaged across all days", main = "Time series plot")

filter(stepsperinterval, steps == max(steps)) %>% select(interval, steps)
```

The 835th 5-minute interval (on average across all the days in the dataset) contains the maximum number of steps (max = 206.1698).


## Imputing missing values

  Calculateing and reporting the total number of missing values in the dataset (the total number of rows with NAs)
```{r}
sum(is.na(data$steps))
```
  I used the mean for that 5-minute interval for filling in all of the missing values in the dataset. Here is creating a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
newsteps <- function(interval) {
    stepsperinterval[stepsperinterval$interval == interval, ]$steps
}
dataFilled <- data  # Make a new dataset with the original data
count = 0  # Count the number of data filled in
for (i in 1:nrow(dataFilled)) {
    if (is.na(dataFilled[i, ]$steps)) {
        dataFilled[i, ]$steps <- newsteps(dataFilled[i, ]$interval)
        count = count + 1
    }
}
cat("Total ", count, "NA values were filled.\n\r")
```

  Making a histogram of the total number of steps taken each day.
```{r}
newperday <- group_by(dataFilled, date)
newstepsperday <- summarise(newperday, steps = sum(steps))
head(newstepsperday)
barplot(newstepsperday$steps, names = newstepsperday$date, xlab = "Date", ylab ="Number of steps", main ="Total number of steps taken per day")
abline(h=mean(newstepsperday$steps), col="red", lwd=2)
```

Calculating and reporting the mean of the total number of steps taken per day. 
```{r}
meanFILLEDstepsperday <- mean(newstepsperday$steps)
meanFILLEDstepsperday
```

 Calculating and reporting the median of the total number of steps taken per day
```{r}
medianFILLEDstepsperday <- median(newstepsperday$steps)
medianFILLEDstepsperday
```

These values differ from the estimates from the first part of the assignment, they are higher. 

The impact of imputing missing data on the estimates of the total daily number of steps is as follows: some days with no values are now filled. You can see the both plots together:

```{r, echo=FALSE, results="asis"}
par(mfrow=c(2,1), mar=c(4,4,3,2))
barplot(stepsperday$steps, names = stepsperday$date, xlab = "Date", ylab ="Number of steps", main ="Total number of steps taken per day OLD")
abline(h=mean(stepsperday$steps), col="red", lwd=2)
barplot(newstepsperday$steps, names = newstepsperday$date, xlab = "Date", ylab ="Number of steps", main ="Total number of steps taken per day NEW")
abline(h=mean(newstepsperday$steps), col="red", lwd=2)
```

## Are there differences in activity patterns between weekdays and weekends?

  Creating a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
WDdataFilled <- dataFilled
WDdataFilled$day <- weekdays(as.Date(WDdataFilled$date))
WDdataFilled2 <- transform(WDdataFilled, day = sub("Monday|Tuesday|Wednesday|Thursday|Friday","Weekday", day))
WDdataFilled3 <- transform(WDdataFilled2, day = sub("Saturday|Sunday","Weekend", day))
```
  Making a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
weekdaydata <- filter(WDdataFilled3, day == "Weekday")
weekdayperinterval <- group_by(weekdaydata, interval)
weekdaystepsperinterval <- summarise(weekdayperinterval, steps = mean(steps))

weekenddata <- filter(WDdataFilled3, day == "Weekend")
weekendperinterval <- group_by(weekenddata, interval)
weekendstepsperinterval <- summarise(weekendperinterval, steps = mean(steps))

par(mfrow=c(2,1), mar=c(4,4,3,2))
plot(weekdaystepsperinterval$interval, weekdaystepsperinterval$steps, type = "l", xlab = "5-minute interval", ylab = "Average number of steps taken", main = "Time series plot WEEKDAYS")
abline(h=mean(weekdaystepsperinterval$steps), col="red", lwd=2)

plot(weekendstepsperinterval$interval, weekendstepsperinterval$steps, type = "l", xlab = "5-minute interval", ylab = "Average number of steps taken", main = "Time series plot WEEKENDS")
abline(h=mean(weekendstepsperinterval$steps), col="red", lwd=2)
```
